---
title: "Baseball_data_analyze"
author: "Yiran Yan"
date: "2022-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

clean the environment
```{r}
rm(list=ls(all=TRUE))
cat("\014")
```

load the library
```{r}
library(ggplot2)
library(here)
library(tidyverse)
```

load the data
*I just use some table that I think usable, you guys can explore the data description and find other usable tables*
```{r}
player <- read.csv(here('baseballdatabank-2022.2',
                        'core','People.csv'))
#str(player)
batting <- read.csv(here('baseballdatabank-2022.2',
                         'core','Batting.csv'))

#str(salary)
pitch <- read.csv(here('baseballdatabank-2022.2',
                       'core','Pitching.csv'))

```


## visualization:
* overview the BMI
**focus on batting table**:
* density of IsoP and BMI
* density of OBP
* density of Hit average
* relationship between a player’s strikeout rate (SO/AB) and his home run rate(HR/AB)

## prediction:
* Variables related to player salary:
* using stepwise to choose variables
* put the choosen variable into nerual network
* compare the rmse of 2 hidden layer and 5 hidden layer
* using olden to figure out which variables are the most affectable to salary



* Correlation between physical condition (BMI standard weight) and  Isolated Power percentage (a measure of slugging ability, Slugging Percentage-Batting average)

SLG = (H + 2B + (2 x 3B) + (3 x HR))/AB
IsoP = (H + 2B + (2 x 3B) + (3 x HR))/AB - H/AB
In US units: BMI = weight (lb) ÷ height2 (inches) * 703

```{r}
options(scipen = 200)

#calculate IsoP
#choose AB>50, yearID > 1995(Updated the competition format in 1995)
batting1 <- batting[batting$AB>50,]

batting1 <- batting1%>%
  group_by(playerID,yearID)%>%
  summarise(AB_sum=sum(AB),
            H_sum=sum(H),
            X2B_sum=sum(X2B),
            X3B_sum=sum(X3B),
            HR_sum=sum(HR))%>%
  mutate(IsoP=(H_sum+X2B_sum+(2*X3B_sum)+(3*HR_sum))/AB_sum-H_sum/AB_sum)

batting1 <- batting1[batting1$IsoP>0,]
batting1 <- batting1[batting1$yearID>1995,]

ggplot(batting1, aes(x=IsoP)) + 
 geom_histogram(aes(y=..density..), colour="light blue", fill="white")+
 geom_density(alpha=.2, color='light blue',fill="#4A77B3",adjust=1.5)

#calculate BMI
player <- player%>%
  group_by(playerID)%>%
  mutate(BMI=weight/(height*height)*703)

#summary(player$BMI)
#str(player$BMI)

ggplot(player, aes(x=BMI)) + 
 geom_histogram(aes(y=..density..), colour="light blue", fill="white")+
 geom_density(alpha=.2, color='light blue',fill="#4A77B3",adjust=1.5)
 
```


```{r}
#combine IsoP and BMI
b_p <- merge(batting1, player, by = "playerID",all=FALSE)#inner join

#check the relationship between BMI and IsoP (using point, bar chart and boxplot)

#point(scatter)
ggplot(b_p,aes(x=BMI,y=IsoP))+
  geom_point()+
  geom_smooth()

#group bmi data
#18.5 and 24.9
b_p<- b_p %>%
  mutate(bmi_group = case_when(
    BMI <= 18.5 ~  'Underweight',
    BMI >18.5 & BMI<=23 ~ 'Normal_1',
    BMI > 23 & BMI <=25 ~ 'Normal_2',
    BMI >25 & BMI <=27 ~ 'Overweight_1',
    BMI > 27 & BMI <= 30 ~'Overweight_2',
    BMI >30 ~ 'Obese'))

#unique(b_p$bmi_group)
#table(complete.cases(b_p$bmi_group))# no na values

#bar chart
mean <-  b_p %>%
  group_by(bmi_group) %>%
  summarize(value = mean(IsoP))
mean %>%
  ggplot(aes(x = fct_reorder(bmi_group, value), y = value)) +
  geom_col(fill='light blue')+
  labs(x='BMI',y='avgerage value of IsoP')
  
#boxplot
ggplot(b_p,aes(x=reorder(bmi_group,IsoP),y=IsoP))+
 geom_boxplot(fill='light blue')+
 labs(x='BMI',y='IsoP')

#correlation
cor(b_p$BMI,b_p$IsoP,method="pearson")
```



#### OBP = (Hits + Walks(BB) + Hit by Pitch) / (At Bats + Walks(BB) + Hit by Pitch + Sacrifice Flies).

```{r}
b_pitch <- merge(batting1, pitch, by = "playerID",all.y=FALSE)
obp <- batting%>%
  group_by(playerID,yearID)%>%
  summarise(hits_sum=sum(H),
            walks_sum=sum(BB),
            HBP_sum=sum(HBP),
            AB_sum=sum(AB),
            SF_sum=sum(SF))%>%
  mutate(OBP=(hits_sum+walks_sum+HBP_sum)/(AB_sum+walks_sum+HBP_sum+SF_sum))
# check the N.A value of obp
table(is.na(obp))
# drop na values
obp <- na.omit(obp)

obp1 <- obp[obp$OBP>0,]
obp1 <- obp1[obp1$yearID>1995,]

ggplot(obp1, aes(x=OBP)) + 
 geom_histogram(aes(y=..density..), colour="light blue", fill="white")+
 geom_density(alpha=.2, color='light blue',fill="#4A77B3",adjust=1.5)+
labs(x='On base percentage')

```

#### Hits average = hits/ab

```{r}
batting2 <- batting[batting$AB>50,] 
 
#Calculate Batting Average (AVG) = Hits/AB 
# <- batting2%>%group_by(playerID)%>%mutate(AVG = H/AB) 
 
batting2 <- batting2%>% 
  group_by(playerID,yearID)%>% 
  summarise(H_sum = sum(H), 
            AB_sum = sum(AB))%>% 
  mutate(AVG = H_sum/AB_sum) 
 
   
# filter out AVG>0 and yearID>1995 
batting2 <- batting2[batting2$AVG>0,] 
batting2 <- batting2[batting2$yearID>1995,]

ggplot(batting2, aes(x=AVG)) +  
 geom_histogram(aes(y=..density..), colour="light blue", fill="white")+ 
 geom_density(alpha=.2, color='light blue',fill="#4A77B3",adjust=1.5)+
 labs(x='Hits Average')

```



```{r}
library(plyr)
dataframe.AB <- ddply(batting, .(playerID), summarize,Career.AB=sum(AB, na.rm=TRUE))
#dataframe.AB
Batting <- merge(batting, dataframe.AB, by="playerID")
Batting.5000 <- subset(Batting, Career.AB >= 5000)
ab.hr.so <- function(d){
  c.AB <- sum(d$AB, na.rm=TRUE)
  c.HR <- sum(d$HR, na.rm=TRUE)
  c.SO <- sum(d$SO, na.rm=TRUE)
  data.frame(AB=c.AB, HR=c.HR, SO=c.SO)
  }
aaron <- subset(Batting.5000, playerID == "aaronha01")
#ab.hr.so(aaron)
d.5000 <- ddply(Batting.5000, .(playerID), ab.hr.so)
d.5000 <- ddply(Batting.5000, .(playerID), ab.hr.so)
ggplot(d.5000,aes(x=HR/AB,y=SO/AB))+
    geom_point()+
    geom_smooth()

```


```{r}
```


```{r}
```

